---
title: "R-Bridge-Final"
author: "A.S"
date: "August 9, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Final Project Description

This project is where you show off your ability to  
(1) use R packages  
(2) change the shape of data in a data frame  
(3) provide basic summary statistics and graphics as part of your exploratory data analysis.

. You should start by installing and loading the R package hflights, which is available on CRAN and described here: https://cran.r-project.org/web/packages/hflights/hflights.pdf

. You should perform some basic transformations on the data frame, such as including only a subset of columns, created one or more new derived columns (such as flight date), and/or filtering out rows.

. You should then create basic summary statistics. You should create at least one interesting question for analysis, and answer the question using two or more graphics in R.

```{r}
#install.packages("hflights")

library(ggplot2)
library(gcookbook)
library(hflights)
library(sqldf)
```

####Get the number of rows in the dataset
```{r}
nrow(hflights)
```

####Get the number of columns in the dataset
```{r}
ncol(hflights)
```

####Get column names in the dataset
```{r}
colnames(hflights)
```

####Check the first 10 rows of the dataset
```{r}
head(hflights, n=10)
```

####Select only required columns
```{r}
hflights_df = hflights[c("UniqueCarrier","Month","DayofMonth", "Year",
                         "Origin","Dest","DepTime", "ArrTime", "DepDelay", "ArrDelay",
                         "Cancelled","CancellationCode")]
head(hflights_df, 10)
```


####Extract column UniqueCarrier into a dataframe carrier_df
```{r}
carrier_df = hflights_df [c("UniqueCarrier")]
```

####Check the number of rows in the data frame carrier_df. Make sure you have all the rows
```{r}
nrow(carrier_df)
```

#### Plot a Bar Graph for each Carrier type. Display Carrier on X-axis and Carrier Count on Y-axis
```{r}
ggplot(carrier_df, aes(x=UniqueCarrier), stat="count") + geom_bar(fill="#AFFAAA", color="black")
```

#### What are the different origins and their count of flights?
```{r}
origins_df=sqldf("select Origin, count(*) As NumberOfFlights from hflights_df group by Origin")
print(origins_df)
ggplot(origins_df, aes(x=Origin, y=NumberOfFlights)) + geom_bar(fill="#9999ff", color="black", stat="identity")
```  

#### What are the top 20 destinations and thier counts of flights?
```{r}
destinations_df = sqldf("select Dest, NumberOfFlights from (select Dest, count(*) As NumberOfFlights from hflights_df group by Dest) order by NumberOfFlights desc")
nrow(destinations_df)
destinations_df = head(destinations_df, 20)
print (destinations_df)
ggplot(destinations_df, aes(x=Dest, y=NumberOfFlights)) + geom_bar(fill="#ff6666", color="black", stat="identity")
```


#### Add a new column called flight date and find the number of flights in each month
```{r}

hflights_df$Date = paste( hflights_df$Month , hflights_df$DayofMonth , hflights_df$Year, sep = "/" )
hflights_df$MonthStr [hflights_df$Month=="1"]  = "01"
hflights_df$MonthStr [hflights_df$Month=="2"]  = "02"
hflights_df$MonthStr [hflights_df$Month=="3"]  = "03"
hflights_df$MonthStr [hflights_df$Month=="4"]  = "04"
hflights_df$MonthStr [hflights_df$Month=="5"]  = "05"
hflights_df$MonthStr [hflights_df$Month=="6"]  = "06"
hflights_df$MonthStr [hflights_df$Month=="7"]  = "07"
hflights_df$MonthStr [hflights_df$Month=="8"]  = "08"
hflights_df$MonthStr [hflights_df$Month=="9"]  = "09"
hflights_df$MonthStr [hflights_df$Month=="10"] = "10"
hflights_df$MonthStr [hflights_df$Month=="11"] = "11"
hflights_df$MonthStr [hflights_df$Month=="12"] = "12"
```

#### Group flights for each month 

```{r}
monthly_df=sqldf("select MonthStr, count(*) As NumberOfFlights from hflights_df group by MonthStr")
print(monthly_df)
ggplot(monthly_df, aes(x=MonthStr, y=NumberOfFlights)) + geom_bar(fill="#8877aa", color="black", stat="identity")

```

#### Find which months have the highest and lowest flights?
```{r}
monthly_df_sorted = sqldf("select MonthStr As Month, NumberOfFlights from monthly_df order by NumberOfFlights desc")
head(monthly_df_sorted, 12)
```

####Confirm that all columns are added 
```{r}
head(hflights_df, n=10)

```

